---
title: "Lab 02"
subtitle: "CS3172-1, Spring 2023, Effat University"
author: "Hanin Alzaher"
output: pdf_document
---

# Packages

```{r load-packages, message = FALSE, warning = FALSE}
library(tidyverse) 
library(scales)
``` 

# Data

```{r read-in-data, message = FALSE}
cas <- read_rds("data/canada_survey.rds")
```

# Tasks

## Task 1

```{r}
cas <- filter(cas, energy_expense>0 , household_income>0)
```

```{r}
cas <- mutate(cas, marital_status = factor(marital_status))
```

## Task 2

```{r}
cas <- mutate(cas, heat_equip= case_when(heat_equip==1~"steam",
                                         heat_equip==2~"forced air",
                                         heat_equip==3~"stove",
                                         heat_equip==4~"electric heating"))
```

```{r}
cas <- mutate(cas, heat_fuel= case_when(heat_fuel==1~"oil",
                                        heat_fuel==2~"gas",
                                        heat_fuel==3~"electricity",
                                        heat_fuel==4~"other"))
```

## Task 3

```{r}
cas %>%
  group_by(heat_fuel, heat_equip)%>%
  summarise( avg = mean(energy_expense),
             median= median(energy_expense),
             standerd_deviation= sd(energy_expense))
  

```

- Provide the answer to the theoretical questions here
-	What combination of fuel type and equipment has the highest average energy expense?
  oil and forced air has the the highest average energy expense with 3498.85
-	Which combination has the most variability with regards to energy expense?
  other and steam with standerd deviation 2278.9
-	Which type of heating equipment doesn’t take all possible fuel types?
  electricity heating.


## Task 4

```{r}
ggplot(data = cas, mapping = aes(x = heat_fuel, y= energy_expense)) +
  geom_boxplot() + facet_wrap(~heat_equip, nrow = 2)+
scale_y_continuous(labels = scales :: dollar_format())+
labs(title = "Energy Expense by Heating Fuel Type",
subtitle = "faceted by type of heating equipment",
x="Heat Fuel Type",
y="Energy Expenses")
```

## Task 5

```{r}
cas <- mutate(cas, prop_energy =
(energy_expense / household_income))

cas2 <- cas %>%
  arrange(desc(prop_energy))%>%
  slice(1,n()) %>% glimpse()

```

## Task 6

```{r message=FALSE}
cas%>%
  group_by(year,province )%>%
  summarise(median_energy_expense_per_room=median(energy_expense/rooms))%>%
  slice_min(median_energy_expense_per_room)
```

## Task 7

```{r fig.width=7, fig.height=6, message=FALSE}
cas %>% mutate( energy_vulnerable= if_else(prop_energy>0.05,"vulnerable","not"))%>%
  group_by(education,age)%>%
  summarise(prop_vul= mean(energy_vulnerable == "vulnerable"))%>%
  ungroup()%>%
  ggplot( aes(x=education,y=age, fill=prop_vul)) +
  geom_raster()  + scale_fill_gradient(low = "white",high = "red") +
  theme_bw() + labs(fill = "Proportion\nenergy\nvulnerable")
```

what we observe from the plot that the older people have vulnerable energy proportion.



